Low Complexity Adaptive Sonar Imaging

### 2016-03-02 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2016:03:02	16:12:30	(on. 2.mars)

To build confidence in the LCA method we show that a robust MVDR implementation typically creates weights spanning between the rectangular and Hamming window function. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the -3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular and Kaiser window with $\beta=5$, in unsteered version, and left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

@@

In this study we expand our previous work of applying LCA to active sonar imaging~\cite{Buskenes2011}. We use variations of the parametric Kaiser window, but also compare it to the trigonometric window function. We extend Synnev\aa{}g's work by steering all our windows and not just some, and demonstrate how LCA's performance is affected by the type and number of windows included. Our results are obtained using experimental data from the 32 element Kongsberg Maritime HISAS1030 sonar, which operates at 100\,kHz with an -3\,dB element opening angle of 23$^\circ$. Although this system was designed as synthetic aperture sonar (SAS) where an image is synthesized from several pings of data, we will only create images out of single pings (sectorscan imaging) in this study.

Our results show that when MVDR is modified to work in active sonar imaging, it computes weights with spatial responses that tend to be symmetric and steered within a fraction of the window's -\,3dB resolution. When letting LCA choose from Kaiser windows with similar responses, we obtain images with improved noise suppression and resolution. The resolution gain is made predictable by limiting the maximum steering to a fraction of the window's -3\,dB width. Overall, LCA produces images that are similar to MVDR, but with a resolution in between that of DAS and MVDR. LCA is also inherently robust, easy to implement, and fairly easy to understand.

@@

\begin{figure}[tbp]\centering%
\ifPeerReview%
\includegraphics[width=.5\linewidth]{gfx/buske1_online.eps}%
\else%
\includegraphics[width=\linewidth]{gfx/buske1_online.eps}%
\fi%
\caption{Each Kaiser window is steered in the interval $\phi=[0, \phi_{\text{-3dB}}(\beta)]$. The angle $\phi_{\text{-3dB}}(\beta)$ is the amount of steering needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.}\label{windows_steering}
\end{figure}

@@

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/buske3_online.pdf}%
\caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, with the pixel groups are indicated with red boxes.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window amplitude and phase responses computed from 40x40 pixel groups in the shadow, highlight and speckle region of the image. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.} }\label{mvdr_selected_windows}
\end{figure*}

\begin{figure*}[t]%
\includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
\caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large and equal size window database, but where the upper bounds for $\beta$ and $\phi$ is varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two -3dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
\newline
\emph{Upper bound $\beta$:}\hfill
\parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR tend to choose windows with levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
\emph{Upper bound $\phi$:}\hfill
\parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
}\label{oversampling_mosaic_bounds}
\end{figure*}

@@

Each window's response is constrained to unit gain in the look direction. Hence, when it is steered the white noise gain must increase and the the signal-to-noise ratio decrease, as shown in \Fig{windows_steering}. To limit this we place an upper bound to the steering, chosen such that wide windows are allowed to be steered more than narrow ones. We call this upper bound for steering $\phi_{\text{3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a \minus{}3\,dB crossing that is exactly half that of the unsteered window. Using a numerical search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values, we obtained a third degree polynomial fit to these points:
%
\begin{align}
\phi_\text{3dB}(\beta) \approx \minus{}0.002894\beta^3 + 0.02451\beta^2 - 0.01308\beta + 0.2980\label{eq:steering_kaiser_3dB}
\end{align}
%
Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

@@

While we focus on using LCA with Kaiser windows, we will for completeness compare the performance with that of using trigonometric windows. This is defined as:
%
\begin{align}
f_m(\alpha) = \alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\label{eq:trig_window_function}
\end{align}
%
where we apply the contraint $\alpha\in[0.5,1]$ to avoid windows with negative coefficients. Steering is applied as in (\ref{eq:steered_window}), and the $\alpha$-value that halves the 3\,dB distance is found using the method described in section \ref{sec:lca_steering_bounds}:
%
\begin{align}
\phi_\text{3dB}(\alpha) \approx \minus{}2.214\beta^3 + 6.084\beta^2 - 5.710\beta + 2.136.
\end{align}

@@

Since LCA operates with the same optimization criterion as MVDR, we will be using approximately a factor 8 lateral oversampling for both beamformers. The images appear more visually pleasing and detailed up to 8. We consider an absolute minimum to be 2, due to the non-linear nature of delaying pixels that isn't strictly in the far field, and of displaying the absolute value of the pixels on a decibel scale. In \mediaI{}\footnote{For media location refer to Appendix \ref{source_code}.} we visualize the effect of changing the lateral oversampling factor on a set of LCA images. Upon display the images are all bilinearly upinterpolated to the same size. 
% 
% \todo{Need $M$ in the image sector. Why???}

@@

To test the performance of the MVDR and LCA beamformers, we have processed data acquired by the Kongsberg Maritime HUGIN AUV carrying their 32 element HISAS1030 sonar~\cite{Hansen2011}. It is a high resolution sonar with 1.2\;m array length, 100\;kHz operating frequency, 30\;kHz bandwidth and 23$^\circ$ element -3\,dB opening angle.

Two different scenes will be studied. One of the 1500 DWT oil tanker wreck Holmengraa lying at a slanted seabed at 77\;m depth outside of Horten, Norway~\cite{holmengraa}. It measures 68\;m by 9\;m and fills most of the highlighted sector. This data was collected by the Norwegian Defense Research Establishment and Kongsberg Maritime. The other scene is of an 1\,m by 1\,m iron resolution test cross, connected to a an anchor with 13\,cm diameter (\Fig{cross}). This data was collected by the Bundeswehr Technical Center for Ships and Naval Weapons, Maritime Technology and Research (WTD 71). Both scenes are imaged in sectorscan mode.

@@

\mediaII{}\footnote{For media location refer to Appendix \ref{source_code}.} animates how \Fig{mvdr_selected_windows} changes as a function of the MVDR subarray length $L$ and temporal averaging $K$. At $L=1$ the window responses are rectangular in all areas in the image. At $L=2$ we observe window responses with slight amplitude variations in the highlight region, but with the phase being 0$^\circ$ in the illuminated seafloor sector. Already at $L=2$ the MVDR is able to greatly suppress noise. This is a common observation; adding a little flexibility to adapt to the scene has a dramatic effect, but further allowing full flexibility is much less significant. The media file also shows that MVDR can be run with subarray sizes $L\in[M/2,5M/8]$, but only with temporal averaging $K=1$ or above.

@@

To determine a suitable range for the Kaiser parameter $\beta$ and steering $\phi$, we constructed large and equally sized window databases containing Kaiser windows with varied upper boundaries for $\beta$ and $\phi$. The lower boundaries were chosen as $\beta=0$ and $\phi=0^\circ$, which includes the rectangularly weighted DAS image. The resulting images are shown in \Fig{oversampling_mosaic_bounds}. In each image we computed the lateral distance $\Delta x$ between the \minus{}3dB points of the anchor, with the reference $\Delta x_\text{ref}$ being the rectangularly weighted DAS image. A fourth order polynomial fit was used in this computation. This measure is also commonly called the full width half maximum (FWHM). In the case of a imaging a point scatter, it is closely related to the resolution of of the system. For a rectangular window, for instance, the system resolution is approximately $\Delta \approx \text{FWHM}/0.89$~\cite{Harris1978}.

The LCA images in \Fig{oversampling_mosaic_bounds} demonstrate the effect of adjusting the upper bound of $\beta$ and $\phi$. As suggested by \Fig{mvdr_selected_windows} we first attempted to use $\beta\in[0,2]$. While this significantly improved sidelobe suppression, a slight further improvement was observed up to $\beta\in[0,5]$. This can likely be explained by the need to compensate for the increased sidelobe levels caused by steering. No noticeable difference were observed for higher values than $\beta=5$. The images also demonstrate that increasing the upper bound for steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution is $\Delta\approx0.0125$ radians (0.72$^\circ$), so a point scatterer imaged at 19\,m range with a rectangular DAS would have a lateral $\text{FWHM}\approx0.125\cdot19\,\text{m}\cdot0.89=21\,\text{cm}$. This is close to the FWHM of the anchor in the DAS image in \Fig{oversampling_mosaic_bounds}, which means that the acoustic fingerprint of the anchor is similar to a point source. This suggests the anchor reflections to be specular and that we view its rounded side. As we increase steering to either 50\% or 100\% of the -3dB width, the FWHM drops to 73\% or 53\% to that of the rectangular window, respectively.

\Fig{oversampling_mosaic_bounds_beta} and \Fig{oversampling_mosaic_bounds_phi} illustrates which $\beta$ and $\phi$ values LCA prefers for different regions in the image of the resolution test cross. Red marker lines are used to pinpoint the cut line, location of the anchor and strong scatters on the cross. Observe that in the anchor region only narrow responses ($\beta=0$) are used, which is the reason for the anchor FWHM being measured to the same value regardless of the upper bound of $\beta$ in \Fig{oversampling_mosaic_bounds}. On other scatterers with strong nearby lateral interference LCA prefers wider windows to suppress the interference. When allowed LCA prefers to steer windows away from the sources, but never exceed the $\phi_\text{3dB}(\beta)$ limit. When exceeding this bound we observed oscillation artifacts in the image.

@@

So far we have found it reasonable to use Kaiser windows in the range $\beta\in[0,5]$, each steered within the range $|\phi|\in[0,\phi_\mathrm{3dB}(\beta)]$. What remains is to determine the number of windows needed. We study this in \Fig{oversampling_mosaic}, where we compared LCA images made from windows databases of different sizes, but with the same parameter boundaries. We included the results from using a single window ($N_\beta=1$) and steering angle ($N_\phi=1$), which corresponds to an unsteered rectangularly weighted DAS.

@@

% \newcommand\cc[1]{\multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf #1}}
% \begin{table}[!b]\centering%\normalsize
% \begin{tabular}[c]{l r r r@{}  r r r}\hline
% \rowcolor{tabBlue} & \cc{Resolution} & \cc{S} & \cc{$\beta$} & \cc{$\varphi^*$} &  \cc{N$_\beta$} & \cc{N$_\varphi$} \\\hline
% & 50\%  & 28 & [0,3] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 50\%  & 28 & [0,6] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 50\%  & 28 & [0,9] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 100\% & 32 &  &:30 \\
% & 150\% & 32 & 1 &:4\\
% \end{tabular}\\
% * Steering is relative to the -3dB width of the window being applied.
% \caption{Qualitative study}\label{parameter_study}
% \end{table}%

@@

% \begin{table}[!b]\centering%\normalsize
% \begin{tabular}[c]{l c c c}\hline
% \rowcolor{tabBlue} & \bf Highlight & \bf Shadow & \bf Speckle  \\\hline
% Symmetry             & Moderate  & High   & High \\
% Mainlobe width       & Narrow    & Wide   & Moderate \\
% Sidelobe suppression & Low       & High   & Moderate \\
% Steering             & Low       & Up to -3dB width & Moderate \\
% Oversampling         & 2x-8x     & 2x-8x  & 2x-8x
% \end{tabular}
% \caption{Summary}\label{tab:summary}
% \end{table}%

@@

We found the LCA to be surprisingly insensitive to the exact window function it is used with, supported by the observation that near identical images were produced whether we used Kaiser windows or trigonometric ones. However, LCA must be able to choose between wide and narrow window responses, since the wide ones offer the best sidelobe suppression and the narrow ones offer the best sensitivity. We achieved good results with 2-3 Kaiser windows in the range $\beta\in[0,5]$, and observed minimal difference when adding more windows or adjusting the upper bound of $\beta$. For the steering angle we suggest a value in the range $\phi\in[\pm\delta\phi_\text{3dB}(\beta)]$, where $\phi_\text{3dB}(\beta)$ is the bandwidth of the respective window, and $\delta$ is a scaling parameter. Setting $\delta$ to either 50\% or 100\% reduces the lateral point target size to 73\% or 53\% compared to the rectangular window, respectively. Hence, this parameter can be used to control the aggressiveness and resolution gain of LCA.

@@

\section{Steering angle}

\begin{figure}[t]%\centering%
\includegraphics[width=\linewidth]{gfx/buske8_online.eps}%
\caption{\emph{Steering angle needed to cut the -3\,dB distance of the Kaiser window's spatial response in half}. The needed amount of steering depends on the number of channels $M$, the element spacing relative to the wavelength $\frac{d}{\lambda}$, and the Kaiser parameter $\beta$. }\label{steering_angle_3dB}
\end{figure}
%
One of the goals of this article was to characterize the windows used by the MVDR method, and then identify a subset of these that were suitable for use with the LCA method. We determined that the Kaiser function could fit the role of producing the relevant windows. The spatial response of a steered Kaiser window depends on the number of channels $M$, the element spacing relative to the wavelength $\frac{d}{\lambda}$, the Kaiser parameter $\beta$, and the amount of steering $\phi$.

To control the resolution gain of LCA, we wanted to determine the extent that the windows need to be steered to cut the angle of the -3\,dB point of the window's spatial amplitude response by a fixed amount, say, by a factor 2 as reference. We computed this steering angle for some common configurations of the mentioned parameters. Refer to \Fig{steering_angle_3dB} for the result. The boundary values of the Kaiser window is shown, i.e. the rectangular window at $\beta=0$ and the near gaussian window at $\beta=5$, as well as one in between at $\beta=2.5$. Observe that for a given $M$ and $\beta$, the angle $\phi_\mathrm{3dB}$ scales near proportionally with $\frac{d}{\lambda}$. We found this to be true also for $\frac{d}{\lambda}\in\{0.5,4\}$. Hence, an approximate figure for a rather wide range of system parameters can be derived off this figure.

For the sake of completeness we also supply the source code, see Appendix \ref{source_code}.

@@

\section{Multimedia Files and Source Code}\label{source_code}

>>>>

To build confidence in the LCA method we show that a robust MVDR implementation typically creates weights spanning between the rectangular and Hamming window function. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the \minus{}3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular and Kaiser window with $\beta=5$, in unsteered version, and left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

@@

In this study we expand our previous work of applying LCA to active sonar imaging~\cite{Buskenes2011}. We use variations of the parametric Kaiser window, but also compare it to the trigonometric window function. We extend Synnev\aa{}g's work by steering all our windows and not just some, and demonstrate how LCA's performance is affected by the type and number of windows included. Our results are obtained using experimental data from the 32 element Kongsberg Maritime HISAS1030 sonar, which operates at 100\,kHz with an \minus{}3\,dB element opening angle of 23$^\circ$. Although this system was designed as synthetic aperture sonar (SAS) where an image is synthesized from several pings of data, we will only create images out of single pings (sectorscan imaging) in this study.

Our results show that when MVDR is modified to work in active sonar imaging, it computes weights with spatial responses that tend to be symmetric and steered within a fraction of the window's \minus{}3\,dB resolution. When letting LCA choose from Kaiser windows with similar responses, we obtain images with improved noise suppression and resolution. The resolution gain is made predictable by limiting the maximum steering to a fraction of the window's \minus{}3\,dB width. Overall, LCA produces images that are similar to MVDR, but with a resolution in between that of DAS and MVDR. LCA is also inherently robust, easy to implement, and fairly easy to understand.

@@

\begin{figure}[tbp]\centering%
\ifPeerReview%
\includegraphics[width=.5\linewidth]{gfx/buske1_online.eps}%
\else%
\includegraphics[width=\linewidth]{gfx/buske1_online.eps}%
\fi%
\caption{Each Kaiser window is steered in the interval $\phi=[0, \phi_{\mathrm{3\,dB}}(\beta)]$. The angle $\phi_{\mathrm{3\,dB}}(\beta)$ is the amount of steering needed for the steered response to have a \protect\minus{}3\,dB crossing that is exactly half that of the unsteered window. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.}\label{windows_steering}
\end{figure}

@@

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/buske3_online.pdf}%
\caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, with the pixel groups are indicated with red boxes.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window amplitude and phase responses computed from 40x40 pixel groups to provide sufficient statistics for the shadow, highlight and speckle region of the image. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.} }\label{mvdr_selected_windows}
\end{figure*}

\begin{figure*}[t]%
\includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
\caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large and equal size window database, but where the upper bounds for $\beta$ and $\phi$ is varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
\newline
\emph{Upper bound $\beta$:}\hfill
\parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR tend to choose windows with levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
\emph{Upper bound $\phi$:}\hfill
\parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
}\label{oversampling_mosaic_bounds}
\end{figure*}

@@

Each window's response is constrained to unit gain in the look direction. Hence, when it is steered the white noise gain must increase and the the signal-to-noise ratio decrease, as shown in \Fig{windows_steering}. To limit this we place an upper bound to the steering, chosen such that wide windows are allowed to be steered more than narrow ones. We call this upper bound for steering $\phi_{\mathrm{3\,dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a \minus{}3\,dB crossing that is exactly half that of the unsteered window. The method we devised for finding it is described in Appendix \ref{appendix:steering_angles}.

Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

@@

While we focus on using LCA with Kaiser windows, we will for completeness compare the performance with that of using trigonometric windows. This is defined as:
%
\begin{align}
f_m(\alpha) = \alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\label{eq:trig_window_function}
\end{align}
%
where we apply the contraint $\alpha\in[0.5,1]$ to avoid windows with negative coefficients. Steering is applied as in (\ref{eq:steered_window}), and the $\alpha$-value that halves the \minus{}3\,dB distance is found using the method described in Appendix \ref{appendix:steering_angles}.

@@

Since LCA operates with the same optimization criterion as MVDR, we will be using approximately a factor 8 lateral oversampling for both beamformers. The images appear more visually pleasing and detailed up to 8. We consider an absolute minimum to be 2, due to the non-linear nature of delaying pixels that isn't strictly in the far field, and of displaying the absolute value of the pixels on a decibel scale. In \mediaI{}\footnote{For media location refer to Appendix \ref{app:source_code}.} we visualize the effect of changing the lateral oversampling factor on a set of LCA images. Upon display the images are all bilinearly upinterpolated to the same size. 
% 
% \todo{Need $M$ in the image sector. Why???}

@@

To test the performance of the MVDR and LCA beamformers, we have processed data acquired by the Kongsberg Maritime HUGIN AUV carrying their 32 element HISAS1030 sonar~\cite{Hansen2011}. It is a high resolution sonar with 1.2\;m array length, 100\;kHz operating frequency, 30\;kHz bandwidth and 23$^\circ$ element \minus{}3\,dB opening angle.

Two different scenes will be studied. One of the 1500 DWT oil tanker wreck Holmengraa lying at a slanted seabed at 77\;m depth outside of Horten, Norway~\cite{holmengraa}. It measures 68\;m by 9\;m and fills most of the highlighted sector. This data was collected by the Norwegian Defense Research Establishment and Kongsberg Maritime. The other scene is of an 1\,m by 1\,m iron resolution test cross, connected to a an anchor with 13\,cm diameter (\Fig{cross}). This data was collected by the Bundeswehr Technical Center for Ships and Naval Weapons, Maritime Technology and Research (WTD 71).

@@

\mediaII{}\footnote{For media location refer to Appendix \ref{app:source_code}.} animates how \Fig{mvdr_selected_windows} changes as a function of the MVDR subarray length $L$ and temporal averaging $K$. At $L=1$ the window responses are rectangular in all areas in the image. At $L=2$ we observe window responses with slight amplitude variations in the highlight region, but with the phase being 0$^\circ$ in the illuminated seafloor sector. Already at $L=2$ the MVDR is able to greatly suppress noise. This is a common observation; adding a little flexibility to adapt to the scene has a dramatic effect, but further allowing full flexibility is much less significant. The media file also shows that MVDR can be run with subarray sizes $L\in[M/2,5M/8]$, but only with temporal averaging $K=1$ or above.

@@

To determine a suitable range for the Kaiser parameter $\beta$ and steering $\phi$, we constructed large and equally sized window databases containing Kaiser windows with varied upper boundaries for $\beta$ and $\phi$. The lower boundaries were chosen as $\beta=0$ and $\phi=0^\circ$, which includes the rectangularly weighted DAS image. The resulting images are shown in \Fig{oversampling_mosaic_bounds}. In each image we computed the lateral distance $\Delta x$ between the \minus{}3\,dB points of the anchor, with the reference $\Delta x_\text{ref}$ being the rectangularly weighted DAS image. A fourth order polynomial fit was used in this computation. This measure is also commonly called the full width half maximum (FWHM). In the case of a imaging a point scatter, it is closely related to the resolution of of the system. For a rectangular window, for instance, the system resolution is approximately $\Delta \approx \text{FWHM}/0.89$~\cite{Harris1978}.

The LCA images in \Fig{oversampling_mosaic_bounds} demonstrate the effect of adjusting the upper bound of $\beta$ and $\phi$. As suggested by \Fig{mvdr_selected_windows} we first attempted to use $\beta\in[0,2]$. While this significantly improved sidelobe suppression, a slight further improvement was observed up to $\beta\in[0,5]$. This can likely be explained by the need to compensate for the increased sidelobe levels caused by steering. No noticeable difference were observed for higher values than $\beta=5$. The images also demonstrate that increasing the upper bound for steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution is $\Delta\approx0.0125$ radians (0.72$^\circ$), so a point scatterer imaged at 19\,m range with a rectangular DAS would have a lateral $\text{FWHM}\approx0.125\cdot19\,\text{m}\cdot0.89=21\,\text{cm}$. This is close to the FWHM of the anchor in the DAS image in \Fig{oversampling_mosaic_bounds}, which means that the acoustic fingerprint of the anchor is similar to a point source. This suggests the anchor reflections to be specular and that we view its rounded side. As we increase steering to either 50\% or 100\% of the \minus{}3\,dB width, the FWHM drops to 73\% or 53\% to that of the rectangular window, respectively.

\Fig{oversampling_mosaic_bounds_beta} and \Fig{oversampling_mosaic_bounds_phi} illustrates which $\beta$ and $\phi$ values LCA prefers for different regions in the image of the resolution test cross. Red marker lines are used to pinpoint the cut line, location of the anchor and strong scatters on the cross. Observe that in the anchor region only narrow responses ($\beta=0$) are used, which is the reason for the anchor FWHM being measured to the same value regardless of the upper bound of $\beta$ in \Fig{oversampling_mosaic_bounds}. On other scatterers with strong nearby lateral interference LCA prefers wider windows to suppress the interference. When allowed LCA prefers to steer windows away from the sources, but never exceed the $\phi_\mathrm{3\,dB}(\beta)$ limit. When exceeding this bound we observed oscillation artifacts in the image.

@@

So far we have found it reasonable to use Kaiser windows in the range $\beta\in[0,5]$, each steered within the range $|\phi|\in[0,\phi_\mathrm{3\,dB}(\beta)]$. What remains is to determine the number of windows needed. We study this in \Fig{oversampling_mosaic}, where we compared LCA images made from windows databases of different sizes, but with the same parameter boundaries. We included the results from using a single window ($N_\beta=1$) and steering angle ($N_\phi=1$), which corresponds to an unsteered rectangularly weighted DAS.

@@

% \newcommand\cc[1]{\multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf #1}}
% \begin{table}[!b]\centering%\normalsize
% \begin{tabular}[c]{l r r r@{}  r r r}\hline
% \rowcolor{tabBlue} & \cc{Resolution} & \cc{S} & \cc{$\beta$} & \cc{$\varphi^*$} &  \cc{N$_\beta$} & \cc{N$_\varphi$} \\\hline
% & 50\%  & 28 & [0,3] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 50\%  & 28 & [0,6] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 50\%  & 28 & [0,9] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 100\% & 32 &  &:30 \\
% & 150\% & 32 & 1 &:4\\
% \end{tabular}\\
% * Steering is relative to the \minus{}3\,dB width of the window being applied.
% \caption{Qualitative study}\label{parameter_study}
% \end{table}%

@@

% \begin{table}[!b]\centering%\normalsize
% \begin{tabular}[c]{l c c c}\hline
% \rowcolor{tabBlue} & \bf Highlight & \bf Shadow & \bf Speckle  \\\hline
% Symmetry             & Moderate  & High   & High \\
% Mainlobe width       & Narrow    & Wide   & Moderate \\
% Sidelobe suppression & Low       & High   & Moderate \\
% Steering             & Low       & Up to \minus{}3\,dB width & Moderate \\
% Oversampling         & 2x-8x     & 2x-8x  & 2x-8x
% \end{tabular}
% \caption{Summary}\label{tab:summary}
% \end{table}%

@@

We found the LCA to be surprisingly insensitive to the exact window function it is used with, supported by the observation that near identical images were produced whether we used Kaiser windows or trigonometric ones. However, LCA must be able to choose between wide and narrow window responses, since the wide ones offer the best sidelobe suppression and the narrow ones offer the best sensitivity. We achieved good results with 2-3 Kaiser windows in the range $\beta\in[0,5]$, and observed minimal difference when adding more windows or adjusting the upper bound of $\beta$. For the steering angle we suggest a value in the range $\phi\in[\pm\delta\phi_\mathrm{3\,dB}(\beta)]$, where $\phi_\mathrm{3\,dB}(\beta)$ is the bandwidth of the respective window, and $\delta$ is a scaling parameter. Setting $\delta$ to either 50\% or 100\% reduces the lateral point target size to 73\% or 53\% compared to the rectangular window, respectively. Hence, this parameter can be used to control the aggressiveness and resolution gain of LCA.

@@

\section{Steering angle}\label{appendix:steering_angles}

\begin{figure}[t]%\centering%
\includegraphics[width=\linewidth]{gfx/buske8_online.eps}%
\caption{\emph{Steering angle needed to cut the \protect\minus{}3\,dB distance of the Kaiser window's spatial response in half}. The needed amount of steering depends on the number of channels $M$, the element spacing relative to the wavelength $\frac{d}{\lambda}$, and the Kaiser parameter $\beta$. }\label{steering_angle_3dB}
\end{figure}
%
One of the goals of this article was to characterize the windows used by the MVDR method, and then identify a subset of these that were suitable for use with the LCA method. We determined that the Kaiser function could fit the role of producing the relevant windows. The spatial response of a steered Kaiser window depends on the number of channels $M$, the element spacing relative to the wavelength $\frac{d}{\lambda}$, the Kaiser parameter $\beta$, and the amount of steering $\phi$.

To control the resolution gain of LCA, we wanted to determine the extent that the windows need to be steered to cut the angle of the \minus{}3\,dB point of the window's spatial amplitude response by a fixed amount, say, by a factor 2 as reference. We computed this steering angle for some common configurations of the mentioned parameters. Refer to \Fig{steering_angle_3dB} for the result. The boundary values of the Kaiser window is shown, i.e. the rectangular window at $\beta=0$ and the near gaussian window at $\beta=5$, as well as one in between at $\beta=2.5$. Observe that for a given $M$ and $\beta$, the angle $\phi_\mathrm{3\,dB}$ scales near proportionally with $\frac{d}{\lambda}$. We found this to be true also for $\frac{d}{\lambda}\in\{0.5,4\}$. Hence, an approximate figure for a rather wide range of system parameters can be derived off this figure.

For the sake of completeness we also supply the source code, see Appendix \ref{app:source_code}.

@@

\section{Multimedia Files and Source Code}\label{app:source_code}
